Automated reading of residual plots with computer vision models

### 2024-01-22 paper/paper.Rmd
put some points down

The algorithm for computing equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution. The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns. So each column of $\boldsymbol{A}$ is a set of realization values of $\boldsymbol{\varepsilon}$. Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual rotation operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$. Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Sliverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$. The KDE computation is done by the `density` function in R.

@@

\noindent where $\boldsymbol{e}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $plot(.)_{h \times w}$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $cv(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance.

@@

@@

@@

@@

Non-normality violations are introduced by defining a non-normal distribution for $\boldsymbol{\varepsilon}$. 
- distribution of x1 and x2
- summary table of all the factor values
- example plots for the violations

@@

>>>>

The algorithm for computing equation \ref{eq:kl-1} starts from simulating $m$ sets of $\boldsymbol{\varepsilon}$ according to the error distribution. The simulated errors are stored in a matrix $\boldsymbol{A}$ with $n$ rows and $m$ columns. So each column of $\boldsymbol{A}$ is a set of realization values of $\boldsymbol{\varepsilon}$. Then, we can get $m$ sets of $\boldsymbol{e}$ stored in the matrix $\boldsymbol{B}$ by applying the residual operator $\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}$. Furthermore, kernel density estimation (KDE) with Gaussian kernel and optimal bandwidth selected by the Sliverman's rule of thumb [@silverman2018density] is applied on each row of $B$ to estimate $p_i(e_i)$ for $i = 1, ..., n$. The KDE computation is done by the `density` function in R.

@@

\noindent where $\boldsymbol{e}$ is a vector of residuals obtained from the fitted model rather than a vector of random variables, $\hat{\boldsymbol{y}}$ is the fitted values also obtained from the fitted model, $plot(.)_{h \times w}$ is a plotting function that generates a residuals vs fitted values plot with fixed aesthetic and then saves it as an image with $h \times w$ pixels and three colour channels, and $cv(.)$ is a computer vision model which takes an $h \times w$ image as input and predicts the distance in the domain $[0, +\infty)$.

@@

Having a computer vision model $cv$ which gives an approximated distance between $0$ and $+\infty$ is not enough to determine if the null hypothesis that the model is correctly specified should be rejected. Theoretically, the distance $D$ for a correctly specified model is $0$, because $P$ will be the same as $Q$, and $log(1 + 0) = 0$. The computer vision model may not necessary predict $0$ for a null plot. Using \ref{fig:false-finding} as an example, it contains a visual pattern which is an indication of heteroskedasticity. We would not expect the model to be able to magically tell if the suspicious pattern is caused by the skewed distribution of the fitted values or the existence of heteroskedasticity. Assume we only train the model with this single image, while $50$% of the time the label is $0$, and $50$% of the time the label is some constant greater than $0$. Then, a perfect model will predict a value between $0$ and that constant to achieve a minimum loss. So, it is expected that our computer vision models will also predict value greater than $0$ for some null plots. It will make more sense if we interpret the model prediction as the visual signal strength of the residual plot. Some null plots could have outliers or strong visual patterns due to randomness, and the model will try to summarise those information into the prediction.

To align with the principle of visual inference, we need to provide a valid null distribution for the test statistic. If we treat the approximated distance $\hat{D}$ as a test statistic, then the null distribution of this statistic can be approximated by the empirical distribution of $\hat{D}$ for a large amount of null plots of given a fitted model. The procedure involves applying the residual rotation technique [@buja2009statistical] on the fitted model to obtain null residuals. The null residuals are then used to make null plots and fed into the model. The results are used to construct an empirical distribution.

@@

With the null distribution of the test statistic,

- How many null samples are needed to get stable 95% quantile
- How do we make decision based on the quantile (reject or not reject)
- p-value calculation (percentage of null vss > vss)

@@

- What is the assumption of bootstrapping
- Any drawback: loss of information
- For each bootstrapped fitted model, we can get a bootstrapped vss and construct a new 95% quantile for the null distribution, and check if H_0 should be rejected
- However, it is very expensive to construct a new 95% quantile for each bootstrapped fitted model
- The 95% quantile constructed using the original fitted model can be considered as an estimate of those new 95% quantiles
- We borrow information from the original fitted model
- We can check how many bootstrapped fitted model should be rejected (like the percentage)
- This gives an overall estimate of how often the assumed regression model are considered to be incorrect if the data can be obtained repetitively from the same data generating process
- If this percentage is relatively large, it gives the analyst an warning that the regression model is inappropriate to the data

@@

Non-normality violations are introduced by defining a non-normal distribution for $\boldsymbol{\varepsilon}$.

- distribution of x1 and x2
- summary table of all the factor values
- example plots for the violations

@@

In order to train a robust computer vision model, we intentionally control the distribution of the target variable $D$ of the training data. Making it uniform between $0$ and $7$. This is accomplished by preparing $50$ buckets each only accepts training samples with $D$  between $[7(i - 1)/49, 7i/49)$ for $i < 50$, where $i$ is the index of the $i$-th bucket. For the $50$-th bucket, any training samples with $D >= 7$ will be accepted. We have prepared 80000 training images, so each bucket can only contain $80000 \div 50 = 1600$ training samples. The simulator will repeatedly sample parameter values from the parameter sample, generate residuals and fitted values using the data generating process, compute the distance, and check if the sample can be accepted by the corresponding bucket, until all the buckets are full.

### 2024-01-22 paper/paper.tex
put some points down

The algorithm for computing equation \ref{eq:kl-1} starts from
simulating \(m\) sets of \(\boldsymbol{\varepsilon}\) according to the
error distribution. The simulated errors are stored in a matrix
\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns. So each column of
\(\boldsymbol{A}\) is a set of realization values of
\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of
\(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by applying
the residual rotation operator
\(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\). Furthermore, kernel
density estimation (KDE) with Gaussian kernel and optimal bandwidth
selected by the Sliverman's rule of thumb \citep{silverman2018density}
is applied on each row of \(B\) to estimate \(p_i(e_i)\) for
\(i = 1, ..., n\). The KDE computation is done by the \texttt{density}
function in R.

@@

\noindent where \(\boldsymbol{e}\) is a vector of residuals obtained
from the fitted model rather than a vector of random variables,
\(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
fitted model, \(plot(.)_{h \times w}\) is a plotting function that
generates a residuals vs fitted values plot with fixed aesthetic and
then saves it as an image with \(h \times w\) pixels and three colour
channels, and \(cv(.)\) is a computer vision model which takes an
\(h \times w\) image as input and predicts the distance.

@@

@@

@@

@@

Non-normality violations are introduced by defining a non-normal
distribution for \(\boldsymbol{\varepsilon}\). - distribution of x1 and
x2 - summary table of all the factor values - example plots for the
violations

@@

>>>>

The algorithm for computing equation \ref{eq:kl-1} starts from
simulating \(m\) sets of \(\boldsymbol{\varepsilon}\) according to the
error distribution. The simulated errors are stored in a matrix
\(\boldsymbol{A}\) with \(n\) rows and \(m\) columns. So each column of
\(\boldsymbol{A}\) is a set of realization values of
\(\boldsymbol{\varepsilon}\). Then, we can get \(m\) sets of
\(\boldsymbol{e}\) stored in the matrix \(\boldsymbol{B}\) by applying
the residual operator \(\boldsymbol{B} = \boldsymbol{R}\boldsymbol{A}\).
Furthermore, kernel density estimation (KDE) with Gaussian kernel and
optimal bandwidth selected by the Sliverman's rule of thumb
\citep{silverman2018density} is applied on each row of \(B\) to estimate
\(p_i(e_i)\) for \(i = 1, ..., n\). The KDE computation is done by the
\texttt{density} function in R.

@@

\noindent where \(\boldsymbol{e}\) is a vector of residuals obtained
from the fitted model rather than a vector of random variables,
\(\hat{\boldsymbol{y}}\) is the fitted values also obtained from the
fitted model, \(plot(.)_{h \times w}\) is a plotting function that
generates a residuals vs fitted values plot with fixed aesthetic and
then saves it as an image with \(h \times w\) pixels and three colour
channels, and \(cv(.)\) is a computer vision model which takes an
\(h \times w\) image as input and predicts the distance in the domain
\([0, +\infty)\).

@@

Having a computer vision model \(cv\) which gives an approximated
distance between \(0\) and \(+\infty\) is not enough to determine if the
null hypothesis that the model is correctly specified should be
rejected. Theoretically, the distance \(D\) for a correctly specified
model is \(0\), because \(P\) will be the same as \(Q\), and
\(log(1 + 0) = 0\). The computer vision model may not necessary predict
\(0\) for a null plot. Using \ref{fig:false-finding} as an example, it
contains a visual pattern which is an indication of heteroskedasticity.
We would not expect the model to be able to magically tell if the
suspicious pattern is caused by the skewed distribution of the fitted
values or the existence of heteroskedasticity. Assume we only train the
model with this single image, while \(50\)\% of the time the label is
\(0\), and \(50\)\% of the time the label is some constant greater than
\(0\). Then, a perfect model will predict a value between \(0\) and that
constant to achieve a minimum loss. So, it is expected that our computer
vision models will also predict value greater than \(0\) for some null
plots. It will make more sense if we interpret the model prediction as
the visual signal strength of the residual plot. Some null plots could
have outliers or strong visual patterns due to randomness, and the model
will try to summarise those information into the prediction.

To align with the principle of visual inference, we need to provide a
valid null distribution for the test statistic. If we treat the
approximated distance \(\hat{D}\) as a test statistic, then the null
distribution of this statistic can be approximated by the empirical
distribution of \(\hat{D}\) for a large amount of null plots of given a
fitted model. The procedure involves applying the residual rotation
technique \citep{buja2009statistical} on the fitted model to obtain null
residuals. The null residuals are then used to make null plots and fed
into the model. The results are used to construct an empirical
distribution.

@@

With the null distribution of the test statistic,

\begin{itemize}
\tightlist
\item
  How many null samples are needed to get stable 95\% quantile
\item
  How do we make decision based on the quantile (reject or not reject)
\item
  p-value calculation (percentage of null vss \textgreater{} vss)
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  What is the assumption of bootstrapping
\item
  Any drawback: loss of information
\item
  For each bootstrapped fitted model, we can get a bootstrapped vss and
  construct a new 95\% quantile for the null distribution, and check if
  H\_0 should be rejected
\item
  However, it is very expensive to construct a new 95\% quantile for
  each bootstrapped fitted model
\item
  The 95\% quantile constructed using the original fitted model can be
  considered as an estimate of those new 95\% quantiles
\item
  We borrow information from the original fitted model
\item
  We can check how many bootstrapped fitted model should be rejected
  (like the percentage)
\item
  This gives an overall estimate of how often the assumed regression
  model are considered to be incorrect if the data can be obtained
  repetitively from the same data generating process
\item
  If this percentage is relatively large, it gives the analyst an
  warning that the regression model is inappropriate to the data
\end{itemize}

@@

Non-normality violations are introduced by defining a non-normal
distribution for \(\boldsymbol{\varepsilon}\).

\begin{itemize}
\tightlist
\item
  distribution of x1 and x2
\item
  summary table of all the factor values
\item
  example plots for the violations
\end{itemize}

@@

In order to train a robust computer vision model, we intentionally
control the distribution of the target variable \(D\) of the training
data. Making it uniform between \(0\) and \(7\). This is accomplished by
preparing \(50\) buckets each only accepts training samples with \(D\)
between \([7(i - 1)/49, 7i/49)\) for \(i < 50\), where \(i\) is the
index of the \(i\)-th bucket. For the \(50\)-th bucket, any training
samples with \(D >= 7\) will be accepted. We have prepared 80000
training images, so each bucket can only contain
\(80000 \div 50 = 1600\) training samples. The simulator will repeatedly
sample parameter values from the parameter sample, generate residuals
and fitted values using the data generating process, compute the
distance, and check if the sample can be accepted by the corresponding
bucket, until all the buckets are full.
